\texttt{SuperScreen}: An open-source package for simulating the magnetic response of two-dimensional superconducting devices

### 2021-11-20 main.tex
Updates from Overleaf

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{examples/images/squid-mutuals.png}
    \caption{Calculating SQUID susceptometer mutual inductance, defined as the fluxoid induced in the SQUID pickup loop per unit current flowing in the field coil: $M=\Phi_\mathrm{PL} / I_\mathrm{FC}$. Top row (a): Schematics of \SuperScreen models for the field coil and pickup loop region of four sizes of scanning SQUID susceptometer (generated using \inline{Device.draw()}) with pickup loop inner radii ranging from $0.1\,\um$ (``Small'') to $3\,\um$ (``X-Large''). The wiring layers BE, W1, and W2 are shown in blue, orange, and green respectively. Middle row (b): Simulated magnetic field $\mu_0H_z$ evaluated at the plane of layer W1, which contains the pickup loop, normalized by the current $I_\mathrm{FC}$ flowing in the field coil (which is located in layer BE). Bottom row (c): Convergence of the four models, defined as the fractional change in mutual inductance between subsequent iterations $i$. For all four models, the simulated mutual inductance, $M_\mathrm{sim}$, falls within the range of mutual inductance values measured in real devices, $M_\mathrm{exp}$, which were reported in Table 1 of Ref.~\cite{Kirtley2016-zz}. The meshes for the ``Small,'' ``Medium,'' and ``Large'' models consisted of approximately 5,000 vertices and 10,000 triangles, whereas the mesh for the ``X-Large'' model consisted of about 7,700 vertices and 15,000 triangles. The two smaller models converge more quickly than the two larger models for reasons discussed in \ref{appendix:numerics}.}
    \label{fig:squid-mutuals}
\end{figure*}

@@

\section{Numerical considerations}
\label{appendix:numerics}
While the numerical method described in Section~\ref{section:implementation} is generally quite robust, it can break down for certain extreme geometries. The stream function $g(x, y)$ represents the local magnetization or density of infinitesimal current loops. The $z$-component of the magnetic field at position $\vec{r}=(x, y, z)$ from a film $F$ with stream function $g$ lying in a plane parallel to the $x-y$ plane at vertical position $z'$ is given by (see Equations~\ref{eq:field_from_kernel} and \ref{eq:kernels}):
\begin{align}
\label{eq:hz-from-kernel}
\begin{split}
    H_z(\vec{r}) &= \int_F Q_z(\vec{r},\vec{r}')g(x', y')\,\mathrm{d}^2r',\,\text{where}\\
    Q_z(\vec{r}, \vec{r}') &=  \frac{2(z-z')^2-\rho^2}
            {4\pi[(z-z')^2+\rho^2]^{5/2}}
\end{split}
\end{align}
and $\rho=\sqrt{(x-x')^2+(y-y')^2}$. Eq.~\ref{eq:hz-from-kernel} is exact for a continuous stream function $g$. However, the discretized version of Eq.~\ref{eq:hz-from-kernel}, in which the double integral over the film area $F$ is replaced by a sum over triangular mesh elements, is only valid if $z-z'$, the vertical distance between the film and the point at which the field is being evaluated, is larger than the typical distance $\delta r$ between vertices in the mesh representing the film. For $z-z'\lesssim\delta r$, the field $H_z(\vec{r})$ resembles that of a discrete set of isolated dipoles located at the mesh vertex positions, rather than that of a continuous sheet of current. This can lead to unphysical results when evaluating the field very close to the surface of a film (for example using \inline{Solution.field_at_position()}), or when solving models involving multi-layer structures where the vertical spacing between layers is much smaller than the lateral extent of the films, in which case the iterative calculation (Section~\ref{section:model:multilayer}) may not converge.

The limitation described above can be seen in the model of the largest SQUID susceptometer described in Section~\ref{section:examples:scanning-squid}, which has a field coil outer radius of $8\,\um$ and a total modeled area of roughly $600\,\um^2$. As shown in Figure~\ref{fig:xlarge-dz-dr}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{examples/images/xlarge-dz-dr.png}
    \caption{\SuperScreen results may be unreliable for multi-layer models where the spacing between layers is smaller than the typical vertex-to-vertex distance, $\delta r$. (a) Field coil - pickup loop mutual inductance for the largest SQUID suscepometer modeled in Section~\ref{section:examples:scanning-squid} as a function of the minimum vertical spacing between layers, $\delta z$. The simulated mutual inductance is shown with blue circles, and a histogram of the mesh vertex-to-vertex distances $\delta r$ is shown in orange (arb. $y$ axis units). When $\delta z\lesssim\delta r$, the model significantly underestimates the mutual inductance. (b) and (c): Out-of-plane magnetic field $\mu_0H_z$ evaluated at the plane of the pickup loop (W1 layer) for (b) the nominal layer spacing with $\delta z<\delta 2$ (indicated with a black $\times$ in (a)) and (c) $\delta z > \delta r$ (indicated with a black $+$ in (a)).}
    \label{fig:xlarge-dz-dr}
\end{figure}

The latter case---mutli-layer structures with closely-spaced layers---is the most challenging class of problem to solve because the iterative method used to solve multi-layer structures (Section~\ref{section:model:multilayer}) is memory-intensive for models with a large mesh (many vertices $p$ and triangles $t$, with typically $p\approx t / 2$) and/or many layers $L$. The average distance between vertices decreases slowly with increasing number of triangles, $\delta r\sim t^{-1/2}$, whereas the size of the (dense floating-point) $p\times t$ matrix that represents the dipole kernel $Q_z(\vec{r},\vec{r}')$ in Eq.~\ref{eq:hz-from-kernel} increases roughly as $t^2$. The end result is that the memory footprint of kernel matrix scales roughly as $\delta r^{-4}$, so decreasing the mean distance between vertices by a factor of 2 increases the memory required by a factor of 16. Furthermore, for a model with $L$ layers there are $\binom{L}{2} = L(L-1)/2$ such kernel matrices needed for each iteration of the calculation outlined in Section~\ref{section:model:multilayer}. In \inline{superscreen.solve()}, these $\binom{L}{2}$ matrices are computed during the first iteration and then cached in memory for use in subsequent iterations. One can force the kernel matrices to be cached to disk if they would otherwise occupy too large a fraction of the available system memory using the \inline{cache_kernel_memory_cutoff} argument, but this comes at a significant performance cost.

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{examples/images/squid-mutuals.png}
    \caption{Calculating SQUID susceptometer mutual inductance, defined as the fluxoid induced in the SQUID pickup loop per unit current flowing in the field coil: $M=\Phi_\mathrm{PL} / I_\mathrm{FC}$. Top row (a): Schematics of \SuperScreen models for the field coil and pickup loop region of four sizes of scanning SQUID susceptometer (generated using \inline{Device.draw()}) with pickup loop inner radii ranging from $0.1\,\um$ (``Small'') to $3\,\um$ (``X-Large''). The wiring layers BE, W1, and W2 are shown in blue, orange, and green respectively. Middle row (b): Simulated magnetic field $\mu_0H_z$ evaluated at the plane of layer W1, which contains the pickup loop, normalized by the current $I_\mathrm{FC}$ flowing in the field coil (which is located in layer BE). Bottom row (c): Convergence of the four models, defined as the fractional change in mutual inductance between subsequent iterations $i$. For all four models, the simulated mutual inductance, $M_\mathrm{sim}$, falls within the range of mutual inductance values measured in real devices, $M_\mathrm{exp}$, which were reported in Table 1 of Ref.~\cite{Kirtley2016-zz}. The meshes for the ``Small,'' ``Medium,'' and ``Large'' models consisted of approximately 5,000 vertices and 10,000 triangles, whereas the mesh for the ``X-Large'' model consisted of about 7,700 vertices and 15,000 triangles. The two smaller models converge more quickly than the two larger models for reasons discussed in \ref{appendix:numerics}. Note that for the ``X-Large'' model, we set the thicknesses of layers I1 and I2 both to 300 nm to ensure convergence (see \ref{appendix:numerics} and Figure~\ref{fig:xlarge-dz-dr}).}
    \label{fig:squid-mutuals}
\end{figure*}

@@

\section{Numerical considerations}
\label{appendix:numerics}
While the numerical method described in Section~\ref{section:implementation} is generally quite robust, it can break down for certain extreme geometries. The stream function $g(x, y)$ represents the local magnetization or density of infinitesimal current loops. The $z$-component of the magnetic field at position $\vec{r}=(x, y, z)$ from a film $F$ with stream function $g$ lying in a plane parallel to the $x-y$ plane at vertical position $z'$ is given by (see Equations~\ref{eq:field_from_kernel} and \ref{eq:kernels}):
\begin{align}
\label{eq:hz-from-kernel}
\begin{split}
    H_z(\vec{r}) &= \int_F Q_z(\vec{r},\vec{r}')g(x', y')\,\mathrm{d}^2r',\,\text{where}\\
    Q_z(\vec{r}, \vec{r}') &=  \frac{2(z-z')^2-\rho^2}
            {4\pi[(z-z')^2+\rho^2]^{5/2}}
\end{split}
\end{align}
and $\rho=\sqrt{(x-x')^2+(y-y')^2}$. Eq.~\ref{eq:hz-from-kernel} is exact for a continuous stream function $g$. However, the discretized version of Eq.~\ref{eq:hz-from-kernel}, in which the double integral over the film area $F$ is replaced by a sum over triangular mesh elements, is only valid if $\delta z = z-z'$, the vertical distance between the film and the point at which the field is being evaluated, is larger than the typical distance $\delta r$ between vertices in the mesh representing the film. For $z-z'=\delta z\lesssim\delta r$, the field $H_z(\vec{r})$ resembles that of a discrete set of isolated dipoles located at the mesh vertex positions, rather than that of a continuous sheet of current (see Figure~\ref{fig:xlarge-dz-dr}(b)). This can lead to unphysical results when evaluating the field very close to the surface of a film (for example using \inline{Solution.field_at_position()}), or when solving models involving multi-layer structures where the vertical spacing between layers is much smaller than the lateral extent of the films, in which case the iterative calculation (Section~\ref{section:model:multilayer}) may not converge.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{examples/images/xlarge-dz-dr.png}
    \caption{\SuperScreen results may be unreliable for multi-layer models where the spacing between layers is smaller than the typical vertex-to-vertex distance, $\delta r$. (a) Field coil - pickup loop mutual inductance for the largest SQUID suscepometer modeled in Section~\ref{section:examples:scanning-squid} as a function of the minimum vertical spacing between layers, $\delta z$. The simulated mutual inductance is shown with blue circles, and a histogram of the mesh vertex-to-vertex distances $\delta r$ is shown in orange (arb. $y$ axis units). When $\delta z\lesssim\delta r$, the model significantly underestimates the mutual inductance. (b) and (c): Out-of-plane magnetic field $\mu_0H_z$ evaluated at the plane of the pickup loop (W1 layer) for (b) the nominal layer spacing with $\delta z<\delta 2$ (indicated with a black $\times$ in (a)) and (c) $\delta z > \delta r$ (indicated with a black $+$ in (a)). For $\delta r\lesssim\delta z$, the magnet field calculated using Eq.~\ref{eq:hz-from-kernel} resembles that of a discrete set of isolated dipoles rather than a continuous sheet of current.}
    \label{fig:xlarge-dz-dr}
\end{figure}

The limitation described above can be seen in the model of the largest SQUID susceptometer described in Section~\ref{section:examples:scanning-squid}, which has a field coil inner radius of $6\,\um$ and a total modeled area of roughly $600\,\um^2$. As shown in Figure~\ref{fig:xlarge-dz-dr}, \SuperScreen significantly underestimates the mutual inductance between the field coil and pickup loop when $\delta z <\delta r$ because, in that case, the discretized version of Eq.~\ref{eq:hz-from-kernel} does not correctly compute the magnetic field due the stream function in the superconducting layers. For the sake of computing mutual inductance as in Section~\ref{section:examples:scanning-squid}, it is physically reasonable to artificially increase the vertical spacing between layers such that $\delta z\geq \delta r$ because we expect the magnetic field a vertical distance $\delta z$ away from the center of the field coil to fall off roughly as $\left(\delta z^2 + R^2_\mathrm{FC}\right)^{-3/2}$, where $R_\mathrm{FC}\approx 6\,\um\gg\delta z$. However, in situations where $\delta z$ is a critical dimension (in the sense that increasing it invalidates the physical model), one's only option is to decrease $\delta r$ by increasing the density of the mesh.

For this reason, mutli-layer structures with closely-spaced layers are the most challenging class of problem to solve. The iterative method used described in Section~\ref{section:model:multilayer} is memory-intensive for models with a large mesh (many vertices $p$ and triangles $t$, with typically $p\approx t / 2$) and/or many layers $L$. The average distance between vertices decreases slowly with increasing number of triangles, $\langle\delta r\rangle\sim t^{-1/2}$, whereas the size of the (dense floating-point) $p\times t$ matrix that represents the dipole kernel $Q_z(\vec{r},\vec{r}')$ in Eq.~\ref{eq:hz-from-kernel} increases roughly as $t^2$. The end result is that the memory footprint of kernel matrix scales roughly as $\langle\delta r\rangle^{-4}$, so decreasing the mean distance between vertices by a factor of 2 increases the memory required by a factor of 16. Furthermore, for a model with $L$ layers there are $\binom{L}{2} = L(L-1)/2$ such kernel matrices needed for each iteration of the calculation outlined in Section~\ref{section:model:multilayer}. In \inline{superscreen.solve()}, these $\binom{L}{2}$ matrices are computed during the first iteration and then cached in memory for use in subsequent iterations. One can force the kernel matrices to be cached to disk if they would otherwise occupy too large a fraction of the available system memory using the \inline{cache_kernel_memory_cutoff} argument, but this comes at a significant performance cost.
